Automated Dynamic Slice Backsolving

### 2014-03-23 docs/paper/fsuthesis/thesis-template/paper.tex
finished testing and graphs

\begin{acknowledgments}
SFS
%Zhi Wang
%Bret Whissel
% Salwan
\end{acknowledgments}

@@

We have implemented this technique in a proof of concept fuzzer, Stubble, which operates on x86 programs. Stubble dynamically instruments binaries, and automatically generates test files which exercise new paths in the input program. Program input, output, and interesting conditions like segfaults are recorded during execution, and taint propagation is recorded from user input sources to branch sinks. When a tainted branch is found, stubble automically backsolves the execution slice and generates a new test case which will drive the program down the opposite path explored in the current execution.

@@

Stubble consists of a pintool which implements DTA, and a Python driver which keeps track of test cases, input and output recorded during execution, and employs dynamic slice backsolving to drive mutation using taint analyses provided by the DTA pintool. Pintools are arbitrary C++ which Pin compiles and inserts into a given binary during instrumentation. The driver accepts a valid x86 binary and a seed test case. From the seed test case, the driver runs the pintool on the binary with the seed test case as input. While the binary runs, the pintool notes when taint in introduced or removed from memory or from registers. If a memory location or a register is tainted, then any operations which move this data elsewhere or use this data in an operation spread taint to the appropriate areas. Moving untainted data into a previously tainted location removes taint from the system. For instance, if the eax register is tainted, then adding eax and ebx and storing the result in ebx will taint the ebx register. However, should an operation move the value of ebx to eax, then eax will no longer be tainted.

@@

    \node [block, left of=Pintool, node distance=4cm] (input) {input/};    
    \node [block, below of=Pintool, below of=input, node distance=2cm] (output) {output/};
    \node [block, right of=output, node distance=4cm] (explored) {explored/};

@@

    % Draw edges
    \path [line] (driver) -- (Pintool);
    \path [line] (Pintool) -- (input);
    \path [line] (Pintool) -- (output);
    \path [line] (Pintool) -- (explored);
    \path [line] (Pintool) -- (decide);
    \path [line] (decide) -- node {yes} (mutate);
    \path [line] (mutate) |- (test case);
    \path [line] (decide) -- node {no}(stop);
    \path [line,dashed] (binary) -- (driver);
    \path [line,dashed] (test case) -- (driver);
    \path [line, dashed] (explored) -- (decide);

@@

@@

(TODO time smaller oracle attacks)
TODO attacks which transform input first to show ability to backsolve

@@

\addplot[
    color=blue,
    mark=square,
    ]
    coordinates {
    		(1,17)
    		(4,20)
    		(7,26)
    		(10,28)
    		(15,38)
    		(20, 45)
    		(25, 53) % 1:41.22elapsed
    		(30, 60) % 1:51.78elapsed
    		(35, 67) % 2:09.49elapsed
    		(40, 74) % 2:19.83elapsed
    		(45, 82) % 2:36.96elapsed
    		(50, 90) % 2:51.77elapsed
    }; 
\end{axis}
\end{tikzpicture}

@@

\begin{figure}
\caption{Secret Guessing Time Elapsed}
\begin{tikzpicture}
\begin{axis}[
    title={Oracle Attack},
    xlabel={Secret Length},
    ylabel={Time Elapsed (seconds)},
    xmin=0, xmax=55,
    ymin=10, ymax=200,
    xtick={0,10,20,30,40,50},
    ytick={0,20,30,40,50,60,80,100,120,140,160,180,200},
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[
    color=red,
    mark=square,
    ]
    coordinates {
    		(1,0) % TODO time these first couple
    		(4,0)
    		(7,0)
    		(10,0)
    		(15,0)
    		(20, 0)
    		(25, 101) 
    		(30, 112)
    		(35, 129) 
    		(40, 140)
    		(45, 157) 
    		(50, 172) 
    }; 
\end{axis}
\end{tikzpicture}
\end{figure}

@@

\begin{biosketch}
Clark Wood is a Scholarship for Service recipient studying computer science at Florida State University. His research and interests focus on information security.
\end{biosketch}

>>>>

\begin{acknowledgments}
%Zhi Wang, for mentoringship, research, paper writing, sounding board

SFS, for the generous funding, mentoring, networking opportunities which have given me the opportunity to study computer science. Specifically my co-principal investigators, Mike Burmester and Xiuwen Liu.

%Bret Whissel
Jonathan Salwan, for his pintool code examples.

\end{acknowledgments}

@@

We have implemented this technique in a proof of concept fuzzer, Stubble, which operates on x86 programs. Stubble dynamically instruments binaries, and automatically generates test files which exercise new paths in the input program. Program input, output, and interesting conditions like segmentation fault signals are recorded during execution, and taint propagation is recorded from user input sources to branch sinks. When a tainted branch is found, stubble automically backsolves the execution slice and generates a new test case which will drive the program down the opposite path explored in the current execution.

@@

Stubble consists of a pintool which implements DTA, and a Python driver which keeps track of test cases, input and output and segfaults recorded during execution, and employs dynamic slice backsolving to drive mutation using taint analyses provided by the DTA pintool. Pintools are arbitrary C++ which Pin compiles and inserts into a given binary during instrumentation. The driver accepts a valid x86 binary and a seed test case. From the seed test case, the driver runs the pintool on the binary with the seed test case as input. While the binary runs, the pintool notes when taint in introduced or removed from memory or from registers. If a memory location or a register is tainted, then any operations which move this data elsewhere or use this data in an operation spread taint to the appropriate areas. Moving untainted data into a previously tainted location removes taint from the system. For instance, if the eax register is tainted, then adding eax and ebx and storing the result in ebx will taint the ebx register. However, should an operation move the value of ebx to eax, then eax will no longer be tainted.

@@

    %\node [block, left of=Pintool, node distance=4cm] (input) {input/};    
    %\node [block, below of=Pintool, below of=input, node distance=2cm] (output) {output/};
    %\node [block, right of=output, node distance=4cm] (explored) {explored/};
	\node [block, left of=Pintool, node distance=4cm] (results) {results};

@@

    % Draw edges
    \path [line] (driver) -- (Pintool);
    %\path [line] (Pintool) -- (input);
    %\path [line] (Pintool) -- (output);
    %\path [line] (Pintool) -- (explored);
	\path [line] (Pintool) -- (results);    
    \path [line] (Pintool) -- (decide);
    \path [line] (decide) -- node {yes} (mutate);
    \path [line] (mutate) |- (test case);
    \path [line] (decide) -- node {no}(stop);
    \path [line,dashed] (binary) -- (driver);
    \path [line,dashed] (test case) -- (driver);
    %\path [line, dashed] (explored) -- (decide);

@@

While finding and executing new code paths is valuable, a fuzzer without a vulnerability condition to monitor for cannot convey meaningful information to an analyst. There are many  ways to detect the presence of a possibly exploitable bug, such as watching for tainted data reaching the instruction pointer, or watching program output for strings that suggest error conditions. Since program output is saved after each run, this is one possible way to detect vulnerabilities with Stubble, however, we specifically wanted to catch segmentation faults, since these are often raised when data corrupts pointers or overflows buffers or stack frames. Stubble monitors for segfaults using the Pin API. A signal-intercepting callback function for signal number 11, segmentation fault, is registered before execution is handed off to the fuzzed program. When a segfault occurs, information about the exception is made recorded and published in the signals/ directory.

@@

TODO attacks which transform input first to show ability to backsolve

@@

\addplot[
    color=blue,
    mark=square,
    ]
    coordinates {
    		(1,17)
    		(5,26)
    		(10,28)
    		(15,38)
    		(20, 45)
    		(25, 53) % 1:41.22elapsed
    		(30, 60) % 1:51.78elapsed
    		(35, 67) % 2:09.49elapsed
    		(40, 74) % 2:19.83elapsed
    		(45, 82) % 2:36.96elapsed
    		(50, 90) % 2:51.77elapsed
    }; 
\end{axis}
\end{tikzpicture}

@@

\begin{figure}
\caption{Secret Guessing Time Elapsed}
\begin{tikzpicture}
\begin{axis}[
    title={Oracle Attack},
    xlabel={Secret Length},
    ylabel={Time Elapsed (seconds)},
    xmin=0, xmax=55,
    ymin=10, ymax=200,
    xtick={0,10,20,30,40,50},
    ytick={0,20,40,60,80,100,120,140,160,180,200},
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[
    color=red,
    mark=square,
    ]
    coordinates {
    		(1,19)
    		(5,30)
    		(10,31)
    		(15,45)
    		(20, 53)
    		(25, 101) 
    		(30, 112)
    		(35, 129) 
    		(40, 140)
    		(45, 157) 
    		(50, 172) 
    };

% TODO find a way to label these

\end{axis}
\end{tikzpicture}
\end{figure}

@@

\begin{biosketch}
Clark Wood is a Scholarship for Service recipient studying computer science at Florida State University. His research interests include reverse engineering, cryptography, networking and their intersection.
\end{biosketch}
